---
title: "Airbnb Bookings Exploratory Data Analasis"
author: "Diana Liang"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd('E:/NYU/2020FALL/Causal/project')

library(dplyr)
library(tidyverse)
library(ggplot2)
library(lubridate)
library(conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

theme_set(theme_bw())
```

```{r}
# read in data
dat <- read_csv('bookings.csv')

# check duplicates
n_distinct(dat$id)==nrow(dat)

# all first active before sign up?
sum(dat$date_first_active<=dat$date_account_created)==nrow(dat)
```
# Treatment: direct marketing or not?
```{r}
dat %>% group_by(treat) %>%
  summarise(n=n()) %>%
  ggplot(aes(factor(ifelse(treat==1,'treat','control')),n,label=n)) +
  geom_bar(stat='identity',width=.4,fill='coral1') +
  labs(x=element_blank(),y='count') +
  geom_text(position=position_dodge(width=1))
```
**More control than treatment.**

# Outcome: Booked or not?
## Overall distribution
```{r}
# overall
dat %>% group_by(is_booked) %>%
  summarise(n=n()) %>%
  ggplot(aes(factor(ifelse(is_booked==1,'booked','not booked')),n,label=n)) +
  geom_bar(stat='identity',width=.4,fill='orange') +
  labs(x=element_blank(),y='count') +
  geom_text(position=position_dodge(width=1))
```

**Overall, more booked than not booked.**
## Distributions within treatment and control
```{r}
# within groups
dat %>% group_by(is_booked,treat) %>%
  summarise(n=n()) %>%
  ggplot(aes(factor(ifelse(treat==1,'treat','control')),n,fill=factor(is_booked),label=n)) +
  geom_bar(stat='identity',position='dodge') +
  theme(legend.title=element_blank()) +
  labs(x=element_blank(),y="count") +
  scale_fill_manual(
    name='Group',
    labels=c('not booked','booked'),
    values=c('lightblue','pink')) +
  geom_text(position=position_dodge(width=1))
```
**In treatment group, booked slightly less than not booked.**
**In control group, more booked than not booked.**

# Covariates
## Gender
```{r}
# overall
dat %>% group_by(gender) %>%
  summarise(n=n()) %>%
  ggplot(aes(gender,n,label=n)) +
  geom_bar(stat='identity',width=.4,fill='orange') +
  labs(x=element_blank(),y='count') +
  geom_text(position=position_dodge(width=1))

# by groups
dat %>% group_by(gender,treat) %>%
  summarise(n=n()) %>%
  ggplot(aes(gender,n,label=n,fill=factor(treat))) +
  geom_bar(stat='identity',position='dodge',width=.4) +
  labs(x=element_blank(),y='count') +
  geom_text(position=position_dodge(width=.5)) +
  theme(legend.title=element_blank()) +
  scale_fill_manual(
    name='Group',
    labels=c('control','treat'),
    values=c('lightblue','pink'))
```
**More female than male, very few other (probably to exclude them?).**
**All gender has good overlap.**
**Within each gender, more control than treat (not well balanced?).**

## Age
```{r}
# overall
dat %>% ggplot(aes(age)) +
  geom_density() 

# age by group
dat %>% ggplot() +
  geom_density(aes(age,fill=as.factor(treat)),alpha=.5) +
  scale_fill_manual(
    labels=c('control','treat'),
    values=c('blue','red')
  ) +
  theme(legend.title=element_blank())
```
**Age is right-skewed.**
**Good overlap: both control and treat cover the whole age range.**
**Balance looks fine for age.**

## English or not?
```{r}
# overall
dat %>% group_by(is_eng) %>%
  summarise(n=n()) %>%
  ggplot(aes(factor(ifelse(is_eng==1,'English','Non-English')),n,label=n)) +
  geom_bar(stat='identity',width=.4,fill='orange') +
  labs(x=element_blank(),y='count') +
  geom_text(position=position_dodge(width=1))

# by groups
dat %>% group_by(is_eng,treat) %>%
  summarise(n=n()) %>%
  ggplot(aes(factor(ifelse(is_eng==1,'English','Non-English')),n,label=n,fill=factor(treat))) +
  geom_bar(stat='identity',position='dodge',width=.4) +
  labs(x=element_blank(),y='count') +
  geom_text(position=position_dodge(width=.5)) +
  theme(legend.title=element_blank()) +
  scale_fill_manual(
    name='Group',
    labels=c('control','treat'),
    values=c('lightblue','pink'))
```
**96% use English as language preference on Airbnb.**
**Good overlap: both treatment and control have English and Non-English.**
**Not perfect balance for language preference.**

## Browser type
```{r}
# overall
dat %>% group_by(browser_group) %>%
  summarise(n=n()) %>%
  ggplot(aes(browser_group,n,label=n)) +
  geom_bar(stat='identity',width=.4,fill='orange') +
  labs(x=element_blank(),y='count') +
  geom_text(position=position_dodge(width=1))

# by groups
dat %>% group_by(browser_group,treat) %>%
  summarise(n=n()) %>%
  ggplot(aes(browser_group,n,label=n,fill=factor(treat))) +
  geom_bar(stat='identity',position='dodge',width=.4) +
  labs(x=element_blank(),y='count') +
  geom_text(position=position_dodge(width=.5)) +
  theme(legend.title=element_blank()) +
  scale_fill_manual(
    name='Group',
    labels=c('control','treat'),
    values=c('lightblue','pink'))
```
**Good overlap and balance for browser type.**

## Signup method
```{r}
# overall
dat %>% group_by(is_mobile) %>%
  summarise(n=n()) %>%
  ggplot(aes(factor(ifelse(is_mobile==1,'mobile','desktop')),n,label=n)) +
  geom_bar(stat='identity',width=.4,fill='orange') +
  labs(x='signup method',y='count') +
  geom_text(position=position_dodge(width=1))

# by groups
dat %>% group_by(is_mobile,treat) %>%
  summarise(n=n()) %>%
  ggplot(aes(factor(ifelse(is_mobile==1,'mobile','desktop')),n,label=n,fill=factor(treat))) +
  geom_bar(stat='identity',position='dodge',width=.4) +
  labs(x='signup method',y='count') +
  geom_text(position=position_dodge(width=.5)) +
  theme(legend.title=element_blank()) +
  scale_fill_manual(
    name='Group',
    labels=c('control','treat'),
    values=c('lightblue','pink'))
```
**Most of the Airbnb users sign up on desktop.**
**Good overlap and balance for signup method.**

## Days to signup
```{r}
# overall
dat %>% ggplot(aes(days_diff)) +
  geom_density() 

# age by group
dat %>% ggplot() +
  geom_density(aes(days_diff,fill=as.factor(treat)),alpha=.5) +
  scale_fill_manual(
    labels=c('control','treat'),
    values=c('blue','red')
  ) +
  theme(legend.title=element_blank())
```
**Assumption: marketing happens after signup behavior.**
**Zero-inflated days to signup: hard to see overlap and balance.**
**Might switch to binary indicator for new comer signup.**

## New-comer signup?
```{r}
dat <- dat %>% mutate(new_signup=ifelse(days_diff==0,1,0))

# overall
dat %>% group_by(new_signup) %>%
  summarise(n=n()) %>%
  ggplot(aes(factor(ifelse(new_signup==1,'New comer','Old visitor')),n,label=n)) +
  geom_bar(stat='identity',width=.4,fill='orange') +
  labs(x='sign up customer type',y='count') +
  geom_text(position=position_dodge(width=1))

# by groups
dat %>% group_by(new_signup,treat) %>%
  summarise(n=n()) %>%
  ggplot(aes(factor(ifelse(new_signup==1,'New comer','Old visitor')),n,label=n,fill=factor(treat))) +
  geom_bar(stat='identity',position='dodge',width=.4) +
  labs(x='sign up customer type',y='count') +
  geom_text(position=position_dodge(width=.5)) +
  theme(legend.title=element_blank()) +
  scale_fill_manual(
    name='Group',
    labels=c('control','treat'),
    values=c('lightblue','pink'))
```
**Too imbalanced to be included?**