---
title: "Airbnb: Causal Inference"
subtitle: "Inferring Causal Effects of Marketing Ads on User Bookings"
author: "Diana Liang, Andrew Pagtakhan"
date: "12/4/2020"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
    number_section: true
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
library(bartCause)
library(knitr)
library(tidyverse)
library(here)
library(gridExtra)
library(rpart)
library(ranger)
library(MatchIt) # remove?
library(arm)
library(gbm)
library(correlationfunnel)
library(conflicted)
library(table1)

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

# set seed
seed <- 14
set.seed(seed)

# increase memory limit for matching functions
# make sure this limit is compatible with the system used to run this code
memory.limit(3.2e10)

knitr::opts_chunk$set(echo = TRUE, 
                      tidy.opts=list(width.cutoff=60), 
                      tidy=TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache.lazy = FALSE)
```

Load data
```{r load_data, message=FALSE, warning=FALSE}
# read in data
dat <- read_csv(here::here('data', 'bookings_sample.csv'))

# convert categorical variables
dat <- dat %>% mutate_at(vars(c("gender","browser_group")),as.factor)

# Airbnb color palette (for plotting)
# Source: https://usbrandcolors.com/airbnb-colors/
col.pal <- c("#FC642D", "#00A699", "#FF5A5F", "#767676", "#484848")
treat.labs <- c("Indirect Ad", "Direct Ad")
treat.name <- "Treatment"
```

# Exploratory Data Analysis and Overlap

There is sufficient overlap for the covariates between those who received 
direct marketing ads vs. indirect ads.

See Tableau workbook for interactive exploration: 
https://public.tableau.com/profile/andrew.pagtakhan#!/vizhome/NYU_Causal_Airbnb/Dashboard?publish=yes


Note: Underlying data used for Tableau workbook is ```bookings_sample.csv```

# Modeling
## Logistic Regression

```{r log_reg, cache = TRUE}
# fit model
fit.lr <- glm(treat~is_eng+is_mobile+age+gender+browser_group,
              family=binomial,
              data=dat)

# extract propensity scores
dat$pscores.lr <- predict(fit.lr,type='response')

## one-to-one nearest neighbor w/ replacement
matches.lr <- matching(z=dat$treat,score=dat$pscores.lr,replace=T)
weight.lr <- ifelse(dat$treat==0,matches.lr$cnts,1)

## IPTW
dat$IPTW.lr <- dat %>% 
  transmute(IPTW=if_else(treat==1,1,pscores.lr/(1-pscores.lr))) %>% 
  unlist() %>% 
  unname()
```


## Probit Regression
```{r prob_reg, cache = TRUE}
# fit model
fit.prob <- glm(treat~is_eng+is_mobile+age+gender+browser_group,
                family=binomial(link='probit'),
                data=dat)

# extract propensity scores
dat$pscores.prob <- predict(fit.prob,type='response')

## one-to-one nearest neighbor w/ replacement
matches.prob <- matching(z=dat$treat,score=dat$pscores.prob,replace=T)
weight.prob <- ifelse(dat$treat==0,matches.prob$cnts,1)

## IPTW
dat$IPTW.prob <- dat %>% 
  transmute(IPTW=ifelse(treat==1,1,pscores.prob/(1-pscores.prob))) %>% 
  unlist() %>% 
  unname()
```

## CART
```{r cart, cache = TRUE}
# fit model
fit.cart <- rpart(treat~is_eng+is_mobile+age+gender+browser_group,
                  data=dat,
                  method='class') # fit a CART model

# extract p scores
# take probability of assigned=1 as propensity scores
dat$pscores.cart <- predict(fit.cart,type='prob')[,2]

## one-to-one nearest neighbor w/ replacement
matches.cart <- matching(z=dat$treat,score=dat$pscores.cart,replace=T)
weight.cart <- ifelse(dat$treat==0,matches.cart$cnts,1)

## IPTW
dat$IPTW.cart <- dat %>% 
  transmute(IPTW=ifelse(treat==1,1,pscores.cart/(1-pscores.cart))) %>% 
  unlist() %>% 
  unname()
```


## Random Forest
```{r rf, cache = TRUE}
fit.rf <- ranger(treat~is_eng+is_mobile+age+gender+browser_group,
                 data=dat)

dat$pscores.rf <- predict(fit.rf,data=dat)$predictions

## one-to-one nearest neighbor w/ replacement
matches.rf <- matching(z=dat$treat,score=dat$pscores.rf,replace=T)
weight.rf <- ifelse(dat$treat==0,matches.rf$cnts,1)

## IPTW
dat$IPTW.rf <- dat %>% 
  transmute(IPTW=ifelse(treat==1,1,pscores.rf/(1-pscores.rf))) %>% 
  unlist() %>% 
  unname()
```

## GBM
```{r gbm, cache = TRUE, message=FALSE}
fit.gbm <- gbm(treat~is_eng+is_mobile+age+gender+browser_group,
               data=dat,
               distribution = "bernoulli")

dat$pscores.gbm <- predict(fit.gbm, type = "response")

## one-to-one nearest neighbor w/ replacement
matches.gbm <- matching(z=dat$treat,score=dat$pscores.gbm,replace=T)
weight.gbm <- ifelse(dat$treat==0,matches.gbm$cnts,1)

## IPTW
dat$IPTW.gbm <- dat %>% 
  transmute(IPTW=ifelse(treat==1,1,pscores.gbm/(1-pscores.gbm))) %>% 
  unlist() %>% 
  unname()
```

## BART

```{r bart_cause, cache = TRUE}
# try bartCause

# get covariates for model
xtrain_bart <- as.data.frame(dat %>% 
                               select(age, gender, is_mobile, 
                                      browser_group, is_eng))
# fit model
bart <- bartc(response = dat$is_booked, 
              treatment = dat$treat,
              confounders = xtrain_bart,
              keepTrees = FALSE,
              method.rsp = "bart",
              method.trt = "bart",
              estimand = "att",
              verbose = FALSE)

# get propensity scores
dat$pscores.bart <- bart$p.score

# get propensity scores
dat$pscores.bart <- bart$p.score

## one-to-one nearest neighbor w/ replacement
matches.bart <- matching(z=dat$treat,score=dat$pscores.bart,replace=T)
weight.bart <- ifelse(dat$treat==0,matches.bart$cnts,1)

## IPTW
dat$IPTW.bart <- dat %>% 
  transmute(IPTW=ifelse(treat==1,1,pscores.bart/(1-pscores.bart))) %>% 
  unlist() %>% 
  unname()
```


# Model Summaries

## Propensity Scores
```{r prop_scores}
# plot propensity scores
dat %>% 
  pivot_longer(cols = contains("pscores")) %>% 
  ggplot(aes(x = value, fill = as.factor(treat))) + 
  geom_histogram(alpha = 0.5) + 
  facet_wrap(facets = vars(name), scales = "free") + 
  labs(title = "Propensity Score Distributions") + 
  theme(legend.position = "bottom") +
  scale_fill_manual(name = treat.name,
                    labels = treat.labs,
                    values = col.pal)
```


## Balance

```{r balance_rev, message=FALSE}
# recode categorical
cat <- dat %>% select(gender,browser_group) %>% binarize()
dat_bal <- bind_cols(dat,cat)

# get covariates and convert to matrices for balance function
bal_mat <- as.matrix(dat_bal %>% select(c(contains(c("browser", "gender")), 
                                      -c("gender", "browser_group","first_browser"),
                                      c("age", "is_mobile", "is_eng"))))
bal_treat <- as.matrix(dat_bal$treat)

# matched weights
match_counts <- list(matches.lr$cnts, dat$IPTW.lr,
                     matches.prob$cnts, dat$IPTW.prob,
                     matches.cart$cnts, dat$IPTW.cart,
                     matches.rf$cnts, dat$IPTW.rf,
                     matches.gbm$cnts, dat$IPTW.gbm,
                     matches.bart$cnts, dat$IPTW.bart)

# convert to matrices
match_mat <- lapply(match_counts, as.matrix)

# compute balance statistics for each model
bal_list <- lapply(match_mat, 
                   function(x) {
                     balance(rawdata = bal_mat, 
                             treat = bal_treat,
                             matched = x,
                             estimand = "ATT")$diff.means.matched
                   })

# model names
model_list <- c('LR 1-1NN','LR IPTW',
                 'PR 1-1NN','PR IPTW',
                 'CART 1-1NN','CART IPTW',
                 'RF 1-1NN','RF IPTW',
                 'GBM 1-1NN','GBM IPTW',
                 'BART 1-1NN','BART IPTW')

# add model name column
bal_dfs <- list()
for (i in 1:length(bal_list)) {
    bal_df <- as.data.frame(bal_list[[i]]) %>% 
      rownames_to_column(var = "Covariate")
    bal_df$model_name <- model_list[[i]]
    bal_dfs[[i]] <- bal_df
}

# add pre-matched balance metrics
bal_df_pre <- as.data.frame(balance(rawdata = bal_mat, 
                                    treat = bal_treat, 
                                    # placeholder - not used
                                    matched = match_counts[[1]], 
                                    estimand = "ATT")$diff.means.raw) %>% 
              rownames_to_column(var = "Covariate")
# add model name
bal_df_pre$model_name <- "prematch"


# combine balance metrics
bal_df <- rbind(bal_df_pre, bind_rows(bal_dfs))

# plot
p.md <- ggplot(bal_df, aes(x = model_name, y = Covariate)) + 
        geom_tile(aes(fill = abs.std.diff)) + 
        labs(title = "Balance: Standardized Difference in Means") + 
        scale_fill_gradient(low = "#ffd9dc", high = "#FF5A5F") + 
        theme(axis.text.x=element_text(angle=45,vjust=.6))

p.sd <- ggplot(bal_df, aes(x = model_name, y = Covariate)) + 
        geom_tile(aes(fill = ratio)) + 
        labs(title = "Balance: SD Ratios") +  
        scale_fill_gradient(low = "#ffd9dc", high = "#FF5A5F") + 
        theme(axis.text.x=element_text(angle=45,vjust=.6))
p.md
p.sd
#grid.arrange(p.md, p.sd, nrow = 2)
```

```{r}
# best model based on difference in means
kable(bal_df %>% 
  group_by(model_name) %>% 
  summarise(mean.bal = mean(abs.std.diff)) %>% 
  arrange(mean.bal), 
  caption = "Absolute Standardized Difference in Means")
```


```{r}
# best model on sd ratios
kable(bal_df %>% filter(Covariate == "age") %>%
    group_by(model_name) %>% 
    summarise(sd.ratio = abs(mean(ratio) - 1)) %>% 
    arrange(sd.ratio), 
    caption = "SD Ratios (Absolute values, centered on 1)")
```


# Estimate Causal Effects

## Treatment Effect for Best Model

Based on the balance metrics, the Probit Regression using IPTW Matching 
was the best model.
```{r}
# logistic regression to get final treatment effect
model_fin <- glm(is_booked ~ is_eng + is_mobile + age + 
                             gender + browser_group + treat, 
              family = quasibinomial(link = "logit"),
              data = dat,
              weights = dat$IPTW.bart)

# odds
odds <- exp(model_fin$coefficients["treat"][[1]])
# probability
paste0("Treatment Effect: ", prob_treat <- odds / (1 + odds))
```

## Treatment effect for all models
```{r}
# treatment effect for all models
effects_list <- lapply(match_counts, 
                       function(x) {
                         glm(is_booked ~ is_eng + is_mobile + age + 
                             gender + browser_group + treat, 
                             family = quasibinomial(link = "logit"),
                             data = dat,
                             weights = x) 
                         }
                       )

# add model name column
eff_dfs <- list()
for (i in 1:length(effects_list)) {
    eff_df <- as.data.frame(effects_list[[i]]$coefficients["treat"])
    eff_df$model_name <- model_list[[i]]
    eff_dfs[[i]] <- eff_df
}

# final treatment effects for each model
eff_fin <- bind_rows(eff_dfs)
colnames(eff_fin) <- c("treat_eff", "model_name")
rownames(eff_fin) <- NULL
kable(eff_fin, caption = "Treatment Effects: All Models")
```

